\propertitle

### 2019-12-09 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-08T20:13:23+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The model can be viewed as a straightforward combination of the
maximum-entropy method and basic sampling relations from the probability
calculus, discussed in \sect~\ref{sec:model}. The maximum-entropy or
minimum-relative-entropy method \citep{jaynes1957}[much clearer
in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b_r1996,grandy1980}
to build initial probability distributions has been used for different
kinds of investigations about the neuronal activity of various brain areas
and about other phenomena of importance to the neurosciences, for example
gene and protein interaction \citep[for
example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
However, our model can also be viewed as an approximation of a more
detailed model within Bayesian theory; we shall constantly keep this point
of view in mind in our discussion.

@@

In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
approximations which define our model, from the point of view of the
probability calculus.

@@

% Later on we'll show that our \dobs\ can be quantified by
% exclusively using the probability calculus. This derivation will provide a
% more accurate quantification, revealing that the maximum-entropy answer is
% only a first approximation.

@@

Step~\ref{item:sample_step} is just an application of the probability
calculus, which gives an exact linear relation between the first $m$
moments for the larger population and the first $m$ for the sample
\citep[\eqns~(16)]{portamanaetal2015}. The ones determine the others and
vice versa at every time bin. This relation holds for any belief
distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
marginal $\pxx(\ya_{t})$ for the sample activity at that bin. These two
distributions are related by
\begin{equation}
  \label{eq:conditional_hypergeometric_t}
  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
  \qquad \ya_{t} \in \set{0,\dotsc,n},
\end{equation}
where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
\enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

One important question remains about the distribution $\px$ of
\eqn~\eqref{eq:app_maxent_pop_first}: \emph{what} is it a distribution of?
-- Of probability? Of relative frequency? The answer is related to an
important relation between maximum-entropy distributions and the
probability calculus, and also to our alternative point of view of the
present model.

@@

Whether such equal beliefs are justified, or for which sampling procedures
they can be justified, is the fundamental, deep question of sampling
theory, which we cannot discuss here \citep[see \eg\ the discussions,
reviews, and references
in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}. It is of
course possible to probabilistically assess, case by case, whether this
equality assumption is appropriate. But we must also be aware that such
assessment in turn rests on analogous assumptions at a higher level, and so
on. The probability calculus, just like the logical calculus, cannot yield
conclusions without premisses \citep[p.~182]{johnson1924}.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-09T10:56:16+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The model can be viewed as a straightforward combination of the
maximum-entropy method and basic sampling relations from the probability-calculus, discussed in \sect~\ref{sec:model}. The maximum-entropy or
minimum-relative-entropy method \citep{jaynes1957}[much clearer
in][]{jaynes1963}{sivia1996_r2006,hobsonetal1973,jaynes1985b_r1996,grandy1980}
to build initial probability distributions has been used for different
kinds of investigations about the neuronal activity of various brain areas
and about other phenomena of importance to the neurosciences, for example
gene and protein interaction \citep[for
example][]{martignonetal1995,bohteetal2000,shlensetal2006,schneidmanetal2006,tkaciketal2006,mackeetal2009b,tkaciketal2009,roudietal2009c,barreiroetal2010,gerwinnetal2010,mackeetal2011b,ganmoretal2011,cohenetal2011,granotatedgietal2013,mackeetal2013,tkaciketal2014b,shimazakietal2015,moraetal2015,lezonetal2006,weigtetal2009}.
However, our model can also be viewed as an approximation of a more
detailed model within Bayesian theory; we shall constantly keep this point
of view in mind in our discussion.

@@

In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
approximations which define our model, from the point of view of the
probability-calculus.

@@

% Later on we'll show that our \dobs\ can be quantified by
% exclusively using the probability-calculus. This derivation will provide a
% more accurate quantification, revealing that the maximum-entropy answer is
% only a first approximation.

@@

Step~\ref{item:sample_step} is just an application of the probability-calculus, which gives an exact linear relation between the first $m$
moments for the larger population and the first $m$ for the sample
\citep[\eqns~(16)]{portamanaetal2015}. The ones determine the others and
vice versa at every time bin. This relation holds for any belief
distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
marginal $\pxx(\ya_{t})$ for the sample activity at that bin. These two
distributions are related by
\begin{equation}
  \label{eq:conditional_hypergeometric_t}
  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
  \qquad \ya_{t} \in \set{0,\dotsc,n},
\end{equation}
where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
\enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

One important question remains about the distribution $\px$ of
\eqn~\eqref{eq:app_maxent_pop_first}: \emph{what} is it a distribution of?
-- Of probability? Of relative frequency? The answer is related to an
important relation between maximum-entropy distributions and the
probability-calculus, and also to our alternative point of view of the
present model.

@@

Whether such equal beliefs are justified, or for which sampling procedures
they can be justified, is the fundamental, deep question of sampling
theory, which we cannot discuss here \citep[see \eg\ the discussions,
reviews, and references
in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}. It is of
course possible to probabilistically assess, case by case, whether this
equality assumption is appropriate. But we must also be aware that such
assessment in turn rests on analogous assumptions at a higher level, and so
on. The probability-calculus, just like the truth-calculus, cannot yield
conclusions without premisses \citep[p.~182]{johnson1924}.
